---
title: "Project"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Library Imports
```{r libraries, warning=FALSE, message=FALSE}
#install.packages("adegenet")
library(adegenet)
library(ggplot2)
library(tidyverse)
library(here)
library(janitor)
library("gapminder")
library(data.table)
library(shiny)
library(shinydashboard)
```

## Import Dataset

```{r datasetImport}
week_gene <- read.csv("data/weeks_genotypes.csv")
head(week_gene)
```

## Determine Heterozygosity

```{r heterozygosity}
alleles <- week_gene[,4:ncol(week_gene)]
genind_alleles <- df2genind(alleles,sep="/",NA.char="NA/NA")
heterozygosity_results <- summary(genind_alleles)
heterozygosity_results
```

## Heterozygosity by Population
```{r popheterozygosity}
pop_geind_alleles <- seppop(df2genind(alleles,sep="/",NA.char="NA/NA", pop=week_gene$Pop))
pop_geind_alleles
```

## Group by Population
```{r popsummary}
MtBuller <- summary(pop_geind_alleles$MtBuller)
MtHiggenbotham <- summary(pop_geind_alleles$MtHigginbotham)
MtBuller
```

## Create dataframes for comparing observed and expected
```{r compare}
MtBullerDf <- data.frame(unlist(MtBuller$Hobs), unlist(MtBuller$Hexp))
names(MtBullerDf) <- c("Hobs", "Hexp")
MtHiggenbothamDf <- data.frame(unlist(MtHiggenbotham$Hobs), unlist(MtHiggenbotham$Hexp))
names(MtHiggenbothamDf) <- c("Hobs", "Hexp")
head(MtBullerDf)
```

## Dataframes Pivoted on Observation Type
```{r pivotCols}
category <- rep(c("Observed", "Expected"), each=length(MtBullerDf$Hobs))
MtBullerPivot <- data.frame(c(seq(1, 24, 1), seq(1, 24, 1)), c(MtBullerDf$Hobs, MtBullerDf$Hexp), category)
MtHiggenbothamPivot <- data.frame(c(seq(1, 24, 1), seq(1, 24, 1)), c(MtHiggenbothamDf$Hobs, MtHiggenbothamDf$Hexp), category)
names(MtBullerPivot) <- c("Location", "Heterozygosity", "Type")
names(MtHiggenbothamPivot) <- c("Location", "Heterozygosity", "Type")
head(MtBullerPivot)
```

## Boxplot Comparison MtBuller
```{r boxplotbuller}
MtBullerPivot %>% 
  ggplot(aes(x=Type, y=Heterozygosity)) + geom_boxplot(na.rm=T) + coord_flip()
```
```{r}
MtBullerPivot %>% 
  ggplot(aes(x=Type, y=Heterozygosity, fill=Type))+
  geom_boxplot(color="grey", alpha=0.8, na.rm = TRUE)+
  geom_violin(alpha=0.5, na.rm = TRUE)+
  scale_fill_brewer(palette = "Pastel1")+
  theme(plot.title = element_text(size = rel(1), face='bold', hjust = 0.5)) +
  labs(title = "MtBuller Observation Type and Heterozygosity",
       x="Type",
       y="Heterozygosity")
```

## Boxplot Comparison MtHiggenbotham
```{r boxplothiggenbotham}
MtHiggenbothamPivot %>% 
  ggplot(aes(x=Type, y=Heterozygosity)) + geom_boxplot(na.rm=T) + coord_flip()
```
```{r}
MtHiggenbothamPivot %>% 
  #mutate(Heterozygosity=as.factor(Heterozygosity)) %>% 
  ggplot(aes(x=Type, y=Heterozygosity, fill=Type))+
  geom_boxplot(color="grey", alpha=0.8, na.rm = TRUE)+
  geom_violin(alpha=0.5, na.rm = TRUE)+
  scale_fill_brewer(palette = "Pastel1")+
  theme(plot.title = element_text(size = rel(1), face='bold', hjust = 0.5)) +
  labs(title = "MtHiggenbotham Observation Type and Heterozygosity",
       x="Type",
       y="Heterozygosity")
```

## Barplot Comparison MtBuller
```{r barplotbuller}
MtBullerPivot %>%
  ggplot(aes(x=Location, y=Heterozygosity, fill=Type)) + geom_bar(stat="identity", position = "identity", alpha=.3, na.rm=T)
```

## Barplot Comparison MtHiggenbotham
```{r barplothiggenbotham}
MtHiggenbothamPivot %>%
  ggplot(aes(x=Location, y=Heterozygosity, fill=Type)) + geom_bar(stat="identity", position = "identity", alpha=.3, na.rm=T)
```

## Scatterplot Comparison of Observed between MtBuller and MtHiggenbotham
```{r scatterobs}
combindedPop <- rbind(MtBullerDf, MtHiggenbothamDf)
combindedPop$Population <- rep(c("MtBuller", "MtHiggenbotham"), each=24)
combindedPop$Location <- c(seq(1, 24, 1), seq(1, 24, 1))
combindedPop %>% 
  ggplot(aes(x=Location, y=Hobs, color=Population)) + geom_point()
```

## Scatterplot Comparison of Expected between MtBuller and MtHiggenbotham
```{r scatterexp}
combindedPop %>% 
  ggplot(aes(x=Location, y=Hexp, color=Population)) + geom_point()
```
```{r}
MtBullerDf %>%
  mutate(Hdiff= Hobs - Hexp) %>%
  arrange(desc(Hdiff))
```

```{r}
MtBullerPivot %>%
  filter(Location %in% c("2","6","1","7","14")) %>%
  ggplot(aes(x=Type, y = Heterozygosity, fill = Type))+ 
  geom_col(alpha=0.8)+
  facet_grid(.~Location)+
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  theme(plot.title = element_text(size = rel(1), face='bold', hjust = 0.5)) +
  labs(title = "Top 5 Microsatellite Loci with Higher than Expected Heterozygosity (MtBuller)",
       x="Type",
       y="Heterozygosity")
```
```{r}
MtHiggenbothamDf %>%
  mutate(Hdiff= Hobs - Hexp) %>%
  arrange(desc(Hdiff))
```
```{r}
MtHiggenbothamPivot %>%
  filter(Location %in% c("2","17","5","9","4")) %>%
  ggplot(aes(x=Type, y = Heterozygosity, fill = Type))+ 
  geom_col(alpha=0.8)+
  facet_grid(.~Location)+
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  theme(plot.title = element_text(size = rel(1), face='bold', hjust = 0.5)) +
  labs(title = "Top 5 Microsatellite Loci with Higher than Expected Heterozygosity (MtHiggenbotham)",
       x="Type",
       y="Heterozygosity")
```

# Most Heterozygous Locus
```{r}
MtBullerDf %>%
  mutate(Heteroz = abs(Hobs - 0.5)) %>%
  arrange(Heteroz)
```
```{r}
MtHiggenbothamDf %>%
  mutate(Heteroz = abs(Hobs - 0.5)) %>%
  arrange(Heteroz)
```
 How many samples do you have from each year?
```{r}
week_gene%>%
  filter(Pop == "MtBuller")%>%
  mutate(Year=as.factor(Year))%>%
  count(Year)
```

Mt Buller Heterozygosity in 2010.
```{r}
week_gene_mtbuller <- filter(week_gene,Pop=="MtBuller")
week_gene_2010 <- filter(week_gene_mtbuller,Year==2010)
alleles_2010 <- week_gene_2010[,4:ncol(week_gene)]
pop_geind_alleles_2010 <- seppop(df2genind(alleles_2010,sep="/",NA.char="NA/NA", pop=week_gene_2010$Pop))
MtBuller_2010 <- summary(pop_geind_alleles_2010$MtBuller)
MtBullerDf_2010 <- data.frame(unlist(MtBuller_2010$Hobs), unlist(MtBuller_2010$Hexp))
names(MtBullerDf_2010) <- c("Hobs", "Hexp")
MtBullerDf_2010 <- tibble::rownames_to_column(MtBullerDf_2010, "Locus")
year2010<-rep(2010,times=24)
MtBullerDf_2010$Year <- year2010
MtBullerDf_2010
```
Mt Buller Heterozygosity in 2011.
```{r}
week_gene_2011 <- filter(week_gene_mtbuller,Year==2011)
alleles_2011 <- week_gene_2011[,4:ncol(week_gene)]
pop_geind_alleles_2011 <- seppop(df2genind(alleles_2011,sep="/",NA.char="NA/NA", pop=week_gene_2011$Pop))
MtBuller_2011 <- summary(pop_geind_alleles_2011$MtBuller)
MtBullerDf_2011 <- data.frame(unlist(MtBuller_2011$Hobs), unlist(MtBuller_2011$Hexp))
names(MtBullerDf_2011) <- c("Hobs", "Hexp")
MtBullerDf_2011 <- tibble::rownames_to_column(MtBullerDf_2011, "Locus")
year2011<-rep(2011,times=24)
MtBullerDf_2011$Year <- year2011
MtBullerDf_2011
```
Mt Buller Heterozygosity in 2012.
```{r}
week_gene_2012 <- filter(week_gene_mtbuller,Year==2012)
alleles_2012 <- week_gene_2012[,4:ncol(week_gene)]
pop_geind_alleles_2012 <- seppop(df2genind(alleles_2012,sep="/",NA.char="NA/NA", pop=week_gene_2012$Pop))
MtBuller_2012 <- summary(pop_geind_alleles_2012$MtBuller)
MtBullerDf_2012 <- data.frame(unlist(MtBuller_2012$Hobs), unlist(MtBuller_2012$Hexp))
names(MtBullerDf_2012) <- c("Hobs", "Hexp")
MtBullerDf_2012 <- tibble::rownames_to_column(MtBullerDf_2012, "Locus")
year2012<-rep(2012,times=24)
MtBullerDf_2012$Year <- year2012
MtBullerDf_2012
```
Mt Buller Heterozygosity in 2013.
```{r}
week_gene_2013 <- filter(week_gene_mtbuller,Year==2013)
alleles_2013 <- week_gene_2013[,4:ncol(week_gene)]
pop_geind_alleles_2013 <- seppop(df2genind(alleles_2013,sep="/",NA.char="NA/NA", pop=week_gene_2013$Pop))
MtBuller_2013 <- summary(pop_geind_alleles_2013$MtBuller)
MtBullerDf_2013 <- data.frame(unlist(MtBuller_2013$Hobs), unlist(MtBuller_2013$Hexp))
names(MtBullerDf_2013) <- c("Hobs", "Hexp")
MtBullerDf_2013 <- tibble::rownames_to_column(MtBullerDf_2013, "Locus")
year2013<-rep(2013,times=24)
MtBullerDf_2013$Year <- year2013
MtBullerDf_2013
```
Mt Buller Heterozygosity in 2014.
```{r}
week_gene_2014 <- filter(week_gene_mtbuller,Year==2014)
alleles_2014 <- week_gene_2014[,4:ncol(week_gene)]
pop_geind_alleles_2014 <- seppop(df2genind(alleles_2014,sep="/",NA.char="NA/NA", pop=week_gene_2014$Pop))
MtBuller_2014 <- summary(pop_geind_alleles_2014$MtBuller)
MtBullerDf_2014 <- data.frame(unlist(MtBuller_2014$Hobs), unlist(MtBuller_2014$Hexp))
names(MtBullerDf_2014) <- c("Hobs", "Hexp")
MtBullerDf_2014 <- tibble::rownames_to_column(MtBullerDf_2014, "Locus")
year2014<-rep(2014,times=24)
MtBullerDf_2014$Year <- year2014
MtBullerDf_2014
```
Mt Buller Heterozygosity in 2015.
```{r}
week_gene_2015 <- filter(week_gene_mtbuller,Year==2015)
alleles_2015 <- week_gene_2015[,4:ncol(week_gene)]
pop_geind_alleles_2015 <- seppop(df2genind(alleles_2015,sep="/",NA.char="NA/NA", pop=week_gene_2015$Pop))
MtBuller_2015 <- summary(pop_geind_alleles_2015$MtBuller)
MtBullerDf_2015 <- data.frame(unlist(MtBuller_2015$Hobs), unlist(MtBuller_2015$Hexp))
names(MtBullerDf_2015) <- c("Hobs", "Hexp")
MtBullerDf_2015 <- tibble::rownames_to_column(MtBullerDf_2015, "Locus")
year2015<-rep(2015,times=24)
MtBullerDf_2015$Year <- year2015
MtBullerDf_2015
```
```{r}
MtBuller_Heterozygosity <- rbind(MtBullerDf_2010,MtBullerDf_2011,MtBullerDf_2012,MtBullerDf_2013,MtBullerDf_2014,MtBullerDf_2015)
MtBuller_Heterozygosity
```

```{r}
library(shiny)

ui <- dashboardPage(
  dashboardHeader(title = "MtBuller Heterozygosity by Locus"),
  dashboardSidebar(disable = T),
  dashboardBody(
    fluidRow(
      box(title = "Plot Options", width = 15,
      selectInput("x", "Select Locus", choices = c("LOC1", "LOC2", "LOC3", "LOC4", "LOC5", "LOC6", "LOC7", "LOC8", "LOC9", "LOC10", "LOC11", "LOC12", "LOC13", "LOC14", "LOC15", "LOC16", "LOC17", "LOC18", "LOC19", "LOC20", "LOC21", "LOC22", "LOC23", "LOC24"), selected = "LOC1"),
    box(title = "MtBuller Heterozygosity", width = 7,
    plotOutput("plot", width = "700px", height = "600px")
    ) 
    ) 
    )) )

 

server <- function(input, output, session) {
  output$plot <- renderPlot({
    MtBuller_Heterozygosity %>% 
      filter(Locus==input$x) %>% 
      ggplot(aes(y=Hobs, x=Year)) +
      geom_col(color="black", fill= "cyan3", alpha=0.6) + 
      theme_light(base_size = 18) + 
      theme(axis.text.x = element_text(angle = 60, hjust = 1)) + 
      labs(x = "Year",
          y = "Heterozygosity")
  })
  session$onSessionEnded(stopApp)
  }


shinyApp(ui, server)
```


 






